---
title: "ProgrammingAssignment3"
author: "Xi Fang"
date: "4/8/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#This data com from the Hospital Compare website which providedata and information about 
#the quality of care at over 4000 Medicare-certified hospitals in the U.S.

1. Plot the 30-day mortality rates for heart attack

```{r check datas, echo=TRUE}
outcomes <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
head(outcomes)
#check the number of columns and rows and names of each column
ncol(outcomes)
nrow(outcomes)
names(outcomes)
#make a histogram of 30day death from heart attack
heartattack <- as.numeric(outcomes[,11])
hist(heartattack)
```

2. Finding the best hospital in a state

```{r besthospital, echo=FALSE}
best <- function(state, outcome) {
  outcomes <- c("heart attack","heart failure", "pneumonia")
  states <- data$State
  ## read outcome data
  data <- read.csv("outcome-of-care-measures.csv")
  colnames(data)[c(11,17,23)] <- c("heart attack", "heart failure", "pneumonia")
  data[,c(11,17,23)] = apply(data[,c(11,17,23)],2, suppressWarnings(as.numeric))
  
  ## check that state and input outcome are valid
  if(outcome %in% outcomes ==FALSE ) {
    stop("invalid outcome")
  }
  if(state %in% states == FALSE) {
    stop("invalid state")
  }
 
  ## Get data in that state and find lowest death for a given outcome
  newdata <- data[data$State==state,]
  ## Remove rows with na in the given outcomes
  comdata<- newdata[complete.cases(newdata[outcome]),]
  ## Return hospital name in that state with lowest 30-day death
  comdata$Hospital.Name[comdata[outcome]== min(comdata[outcome])]
}

best("TX","heart attack")
best("SC", "heart attack")
best("NY", "pneumonia")
best("AK", "pneumonia")

```

```{r besthospital, echo=FALSE}
best <- function(state, outcome) {
  ## read outcome data
  data <- read.csv("outcome-of-care-measures.csv")
  outcomes <- c("heart attack","heart failure", "pneumonia")
  states <- data$State
  colnames(data)[c(11,17,23)] <- c("heart attack", "heart failure", "pneumonia")
  data[,c(11,17,23)] = apply(data[,c(11,17,23)],2, suppressWarnings(as.numeric))
  
  ## check that state and input outcome are valid
  if(outcome %in% outcomes ==FALSE ) {
    stop("invalid outcome")
  }
  if(state %in% states == FALSE) {
    stop("invalid state")
  }
  ## Get data in that state and exclude rows with na in the given outcome
  newdata <- data[data$State==state & !is.na(data[outcome]),]
  ## Return hospital name in that state with lowest 30-day death of the given outcome
  newdata$Hospital.Name[newdata[outcome]== min(newdata[outcome])]
  
}

# best("AL","heart attack")
# best("TX","heart attack")
# best("TX", "heart failure")
# best("MD", "heart attack")
```

3. Ranking hospitals by outcome in a state
```{r ranking, echo=TRUE}
rankhospital <- function(state, outcome, num="best"){
  #Read outcome data
  data <- read.csv("outcome-of-care-measures.csv")
  outcomes <- c("heart attack","heart failure", "pneumonia")
  states <- data$State
  colnames(data)[c(11,17,23)] <- c("heart attack", "heart failure", "pneumonia")
  data[,c(11,17,23)] = apply(data[,c(11,17,23)],2, suppressWarnings(as.numeric))
  data[,c(2,7)] = apply(data[,c(2,7)],2, suppressWarnings(as.character))
  
  ## check that state and input outcome are valid
  if(outcome %in% outcomes ==FALSE ) {
    stop("invalid outcome")
  }
  if(state %in% states == FALSE) {
    stop("invalid state")
  }
  ## Get data in that state and exclude rows with na in the given outcome
  newdata <- data[data$State==state & !is.na(data[outcome]),]
  ## Rank hospitals based on the name by alphabetic 
  rankname <- newdata[order(newdata$Hospital.Name, decreasing = FALSE),]
  ## Rank hospitals based on a given outcome
  rankoutcome <- rankname[order(rankname[outcome],decreasing = FALSE),]
  rownumber <- nrow(rankoutcome)
  ## Get the name of the selected number of ranking
  if (num == "best") {
    newdata$Hospital.Name[newdata[outcome]== min(newdata[outcome])]
  } else if (num == "worst") {
    newdata$Hospital.Name[newdata[outcome]== max(newdata[outcome])]
  }
  else if (num <=rownumber) {
  rankoutcome[num,]$Hospital.Name
  }
  else {
    "NA"
  }
  
}

#rankhospital("TX", "heart failure",4)
#rankhospital("MD", "heart attack","worst")
#rankhospital("MN", "heart attack",5000)
#rankhospital("NC", "heart attack", "worst")
#rankhospital("WA", "heart attack", 7)
rankhospital("TX", "pneumonia", 10)
rankhospital("NY", "heart attack", 7)
```
4. Ranking hospitals in all states
```{r rankall, echo=TRUE}
rankall <- function(outcome, num="best") {
  #Read outcome data
  data <- read.csv("outcome-of-care-measures.csv")
  outcomes <- c("heart attack","heart failure", "pneumonia")
  data[,c(11,17,23)] = apply(data[,c(11,17,23)],2, suppressWarnings(as.numeric))
  data[,2]<-as.character(data[,2])
  ## check input outcome are valid
  if(outcome %in% outcomes ==FALSE ) {
    stop("invalid outcome")
  }
 
 ## Find hospital of a given rank for each state
  col <- if (outcome=="heart attack"){
    11
  }else if (outcome=="heart failure"){
    17
  } else {
    23
  }
  
 # data[,col] <- suppressWarnings(as.numeric(levels(data[,col])[data[,col]]))
   # generate an empty vector that can be filled row by row
   output <- vector()
   states <- levels(data[,7])
  for(i in 1:length(states)) {
    statedata <- data[grep(states[i], data$State), ]
    orderdata <- statedata[order(statedata[, col],statedata[, 2],
                                 na.last = NA), ]
    ## Get the name of the selected number of ranking
  hospital <- if(num == "best") {
    orderdata[1,2]
  } else if(num == "worst") {
    orderdata[nrow(orderdata),2]
  }else{
  orderdata[num,2]
  }
  output <- append(output,c(hospital,states[i]))
}
  ## Return a data frame with the hospital names and state abbr
  output <- as.data.frame(matrix(output, length(states),2,byrow = TRUE))
  colnames(output) <- c("hospital","state")
  rownames(output) <- states
  output
}

#head(rankall("heart attack", 20), 10)
#tail(rankall("pneumonia","worst"),3)
#tail(rankall("heart failure"),10)
#r <- rankall("heart attack", 4)
#as.character(subset(r, state == "HI")$hospital)
#r <- rankall("pneumonia", "worst")
#as.character(subset(r, state == "NJ")$hospital)
r <- rankall("heart failure", 10)
as.character(subset(r, state == "NV")$hospital)

```



































Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
